Type-Safe Modular Parsing

### 2016-11-08 paper/sections/CaseStudy.tex
Finish the script
Fix, add table header
Text

\begin{table}
    \centering
    \input{resources/CaseStudyTable.tex}
    \caption{Comparison of SLOC and execution time}
    \label{tab:comparison}
\end{table}

Table \ref{tab:comparison} shows data of the comparison. The overall result of SLOC comparison is that 60\% of code in total is reduced using our library. That is because the code of some common syntax structures such as variable, lambda abstraction, etc., is reused several times in the whole project. We can see that the first two calculi \lstinline{Arith} and \lstinline{Untyped} we do worse than Ilya's implementation, because in such two cases we do not reuse any existing components. But in each of the following 16 calculi, we can compose our syntax definition, parser and pretty printer from extending reusable components, resulting considerably code reduction.

>>>>

\begin{table}
    \centering
    \begin{tabular}{|l|r|r|r|r|r|r|}
      \hline
        \multirow{2}{*}{\bfseries Calculus Name} & \multicolumn{3}{ c| }{\bfseries SLOC} & \multicolumn{3}{ c| }{\bfseries Time (ms)} \\ \cline{2-7}
        \multicolumn{1}{|c|}{} & \bfseries Ilya's & \bfseries Ours & \bfseries (+/-)\% & \bfseries Ilya's & \bfseries Ours & \bfseries (+/-)\% \\
      \hline
      \input{resources/CaseStudyTable.tex}
      \hline
    \end{tabular}
    \caption{Comparison of SLOC and execution time}
    \label{tab:comparison}
\end{table}

Table \ref{tab:comparison} shows data of the comparison. The overall result of SLOC comparison is that 66.3\% of code is reduced using our library. That is because the code of some common syntax structures such as variable, lambda abstraction, etc., is reused several times in the whole project. We can see that the first two calculi \lstinline{Arith} and \lstinline{Untyped} we do worse than Ilya's implementation, because in such two cases we do not reuse any existing components. But in each of the following 16 calculi, we can compose our syntax definition, parser and pretty printer from extending reusable components, resulting considerably code reduction.

For the execution time comparison, we built a generator to randomly generate valid expressions according to the syntax of each calculus. Each test file corresponds with one calculus and it consists of 500 expressions randomly generated. The size of test files varies from 20KB to 100KB. After testing both implementation, we found that our version is about \huang{60\%} slower. Part of the reason is that our library requires more intermediate function calls, and the object algebra and self-reference needed to be passed as arguments every time. However, another important reason is that, because of the modular setting, we use a different alternative combinator which will try both branches and match the longest one. \huang{ref to somewhere before?} In contrast Ilya uses the ordinary alternative combinator which tries the second branch only if the first one fails. Although this difference slows our version, we want to mention that because of the memorization of Packrat parsers, we are just constant times slower, the overall complexity is still polynomial regarding to the length of the input.

### 2016-11-08 paper/sections/Overview.tex
Text

>>>>

Our modular parsing library consists of four parts, and we will go through them in this section. The first is underlying parsing technique we used. Parsing has been heavily studied over the years, especially for context-free grammars and its subsets. However, the choice of parsing techniques is not a minor issue here, because of modularity. To achieve modularity, the parsers must be written in a way that can be composed, and different parsing techniques have different suitability for that.

Although every parsing algorithm is able to implement by hand in principle, some of them are not friendly for that. A parser generator is a tool to generate parsers automatically or semi-automatically, from a grammar given by the user. There is no restriction on the algorithm, however most of them adopts LL and LR parsing algorithms, which are table-based.

\huang{TALK ABOUT PREVIOUS WORKS USING PARSER GENERATORS, AND ALSO THE DISADVTANGES OF THEM}

Comparing with the parser generators, a parser combinator takes several parsers and produce a new parser as its output. It is popular in functional programming, where the parsers are represented by functions and parser combinators are higher-order functions accepts them. Parser combinators are very suitable for our purpose, because of two reasons. Firstly, they are naturally modular. The manner of using them is to write small parsers and use combinators to composed them together. The construction procedure is explicit and fully controlled by the programmer. Secondly, each parser combinator is represented by a piece of code, and also are the parsers it takes. Thus in a statically typed programming language they can be type-checked during compile time.

There are variants in the category of parser combinators. Almost all of them adopt top-down, recursive descent parsing, which uses backtracking to search through the possible branches. Simple backtracking parser combinators, such as the famous library Parsec in Haskell, have some drawbacks. One is that they cannot support left-recursive grammars. The common solution is to rewrite a left-recursive grammar into an equivalent one, so called left-recursion elimination. This solution requires the whole grammar, so that it cannot be applied in a modular setting because we only have parts of the grammar. Another drawback is performance.
